---
title: "On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  html_document:
    number_sections: yes
    theme: flatly
csl: new-phytologist.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
^2^ URFM, INRA, 84000, Avignon, France
^3^ Université Clermont-Auvergne, INRA, PIAF, 63000, Clermont-Ferrand, France


\
**Corresponding author**: 
Remko Duursma

\
E: remkoduursma@gmail.com

\
T: +61(0)45701806


```{r include=FALSE}
includefigs <- FALSE
library(png)
library(grid)
library(knitr)
library(pander)

source("R/load.R")
```


**Contents:**

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from cuvette data
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to drought
\t4. Acclimation and instantaneous response to temperature
\t5. Other factors
VIII. Conclusions", header=F)

knitr::kable(tab, col.names="")


```


# Notes

@fernandez2017 : more discussion of cuticular composition, increase of gmin with altitude ('unripe' cuticles due to short growing season), foliar uptake of water (and adaptation of the cuticle to facilitate that)

Relationship between gmin and yield (to mention in the text???):
Clarke et al. 1989 . Genotypes with low gmin yielded 40-150% more in 4 out of 5 driest enviroments in 9 years of study.
Clarke and McCaig (1982 ) low order but significant positive correlation between yield and water retention (low gmin) in genotypes of Triticum

Perhaps include here changes with ontogeny is species with different juvenile and adult leaves?
Climent et al. 2006  and Pardos et al. 2009 - juvenile needles in some pines have higher rate of water loss than adult needles of the same individuals. Darrow et al. 2002  found also that in some New Zealand heteroblastic sp juveniles had a higher gmin than adults but other sp no significant differences.




# Summary

Stomata exert a strong control on the transpiration rate of leaves. During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.


# I. Introduction

\
Stomata exert a strong control on the transpiration rate of leaves [@bange1953]. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. There is increasing recognition that the minimum conductance plays an important role in estimating water fluxes in plant canopies [@barnard2013], during heat waves [@kala2016], and in models of plant drought response [@blackman2016; @martinstpaul2017].

\
Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq. 1).

\
(1)
$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions and measurements compare quantitatively, whether it is feasible to assign species-specific values of the minimum conductance, and what are the ecological and environmental controls on g~min~ [@schuster2017].

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research. Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate that the large acclimatory potential of this trait to environmental drivers. 



# II. Comparison of various definitions of minimum conductance

**Table 1.** Definitions of minimum conductance.

| Variable      |  Definition                                                                                       |
| ------------- | ------------------------------------------------------------------------------------------------- |
| g~cut~        | Conductance of an isolated non-stomatal cuticle.                                                  |
| g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
| g~dark~       | Night-time conductance, or conductance after significant dark adaptation.                         |
| g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~    | Cuvette measurements of g~s~ when photosynthesis rates are very low.                              |

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss cuticular conductance in Section V. 

\
The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved. From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. 

\ 
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.  Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^ was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.

\
The fitted minimum conductance parameter, g~0~ (mean = 32.8 mmol m^-2^ s^-1^), was much higher than g~min~, but similar to g~dark~. However, we point out the uncertainties associated with g~0~ in Section III, which tend to inflate estimates of g~0~. It is telling that actual measured stomatal conductance during low photosynthesis (g~s,low A~) was  much lower than g~0~ or g~dark~. In other words, actual daytime conductance during these conditions that cause low photosynthetic rates is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their Fig. A2). 


```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but some data occurs outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cut,isol~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.


# III. Problems with estimating minimum conductance from cuvette data

\
The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.

\
It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated. Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves. Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the slope was unaffected).

\
In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated. We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied. In addition, it is difficult to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study. The correlation between slopes and intercepts is not unique to Eq. 1, but a general property of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry). In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because environmental conditions varied little. We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq. 1 is lower (Fig. 3c). 


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure3.png")
```

**Figure 3**. Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases. (a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.  Vertical lines are 95% confidence intervals. The grey line is a fitted loess smoother with 95% confidence interval. Note the wide confidence intervals, and frequent negative values. (b) Similar to (a), but using the published compilation by @miner2017. The grey line is a fitted loess smoother with 95% confidence interval. (c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e. right-hand side of the equation being fitted) is lower. 



# IV. Use of minimum conductance in models

## 1. Models of water-use efficiency

\
In the Ball-Berry type framework of stomatal conductance models, the effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers. Here we show that the g~0~ parameter not only sets a minimum value of g~s~ in the model (when A~n~ is zero), it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake, @urban2017].

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD). When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

\
(2)
$$A_n/g_s = {C_a}f(D) / g_1$$

and using the diffusion constraint $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,

\
(3)
$$C_i / C_a = 1 - \frac{f(D)}{1.6 \cdot g_1}$$

Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1. 


```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

**Figure 4**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].



## 2. Models of plant desiccation

\
When plants are sufficiently water-stressed that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 5).


```{r fig.height=6, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**. Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants. The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues. (a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 1.6 or 3.3 mmol m^-2^ s^-1^ - all other parameters were equal. (b) Water potential in the soil and leaf as the dry-down progresses. Note the brief increase in soil water potential around days 40 - 70; this represents water released from stem cavitation. (c) Progression of percent loss conductivity (PLC) of the xylem. If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer (or nearly 40%).



# V. Cuticular conductance

\
A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section VII).

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle? 

\
It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017] The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). As such, thick cuticles may be advantageous because prolonged exposure to wind increases the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation explains only some of the wide range of g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight. 


# VI. Nighttime conductance

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition where no photosynthesis takes place. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to any other measure of minimum conductance. The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.


# VII. Environmental and ecological variation in minimum conductance 

## 1. Contribution of stomata to minimum conductance

A few studies have directly quantified the stomatal component of g~min~. A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky. However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.


## 2. Minimum conductance shows little meaningful interspecific variation

\
In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants. We also further analyze the literature compilation of g~min~ (Fig. 6). Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature (all extracted across the range of the species) and g~min~, nor were there meaningful relationships with other traits (**not shown**). We thus found it difficult to explain the variation in g~min~ between species. A simple breakdown by taxonomic order (Fig. 6b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values. 

\
Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories. Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ / g~cut~ by plant growth form. One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~. Such correlation did not bear out across all species in our database, and interesting to note is that *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig. 6b).

\
**crops**


```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Analysis of a literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. (a) Histogram (probability density) of all estimates (after averaging by species, n=`r nrow(gmindat)`), with a log-normal distribution curve (mean = `r round(exp(mean(log(gmindat$gmin))),2)`, sd = `r round(exp(sd(log(gmindat$gmin))),2)`). (b) g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95/% confidence intervals. Numbers above the figure refer to the number of species. Different letters denote significant differences (at $\alpha$ = 0.05, adjusted for multiple comparisons), and grey symbols are species-level data. (c) g~min~ estimates for crops only, averaged by genotype. Bars denote the range, illustrating the wide range in g~min~ between genotypes for a particular crop species. Numbers above the figure refer to the number of genotypes included.



## 3. Acclimation to drought

\
From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a significant  decrease in g~min~ (see Table S2 for a quantitative summary of the literature), with just one exception. The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70%; no attempt at a meta analysis was made since no drought is created equal. In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978]. We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig. 7).

\
A number of studies have attempted to attribute this drought acclimation in g~min~ to a change in the chemical composition of the cuticle. @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes. However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability. @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts). @bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way. Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to g~min~.

\
atmospheric drought
@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content. This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss.


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure7.png")
```

**Figure 7**. Minimum conductance (g~min~) measured by mass loss of detached leaves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or subjected to long-term (8 months) mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N), and ordered by g~min~ in the  well-watered treatment. For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).



## 4. Acclimation and instantaneous response to temperature



The response of minimum conductance to temperature is also striking; it shows both an instantaneous response to increased temperature, as well as an acclimatory response to growth temperatures. @riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures. 

@schuster2016a reported on a correlation between thermal tolerance and g~min~ relationship across nine species. 

Also see @paje1988 for more complex response to temperature (figure out what he really showed).


Lower g~min~ in the trees acclimated to high temperature is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time. Note that this response is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001). 

```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure8.png")
```

**Figure 8**. Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C) [see @drake]. Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There was no consistent effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, mixed-effects model).



## 5. Increased minimum conductance with leaf age


@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age. See SuppInfo 3 for a summary of six studies on leaf age effects. All but one species show an increase in g~min~ with leaf age. Also discussed above in Section~(cuticular conductance), the cuticle is constantly exposed to physical (wind, rain) and biotic (insects, fungi) agents which damage the cuticle over time. 

- do any of those papers say why? summarize

@anfodillo2002 gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude [e.g. @delucia1984; @herrick1991]. @fernandez2017 summarized the literature on this topic and argue that the short growing seasons at high altitude cause 'unripe' cuticles.




# VIII. Conclusions

In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression, the very high values compared to minimum conductance reached for detached leaves, we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.

Then, how should g~0~ be estimated? It is clear from our comparisons that there is no single minimum conductance. Leaves maintain much higher g~dark~ (itself an actively controlled process) than the minimum obtained for detached leaves. Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants. Likewise, as briefly mentioned in the introduction, yet a different g~0~ is reached when temperature is so high that photosynthesis does not take place. It is not our intention to make models more complex, but certainly we cannot avoid the conclusion that g~0~ should be replaced with a more complicated expression.

\
It is probably unavoidable to improve the Ball-Berry model with an expression like,

$$ g_s = g_{min} + (g_0 - g_{min}) + g_1 \frac{A_n}{C_a}f(D)$$

where g~min~ is estimated via MLD, representing a true whole-leaf minimum that is reached under drought conditions, and g~0~ is the minimum reached in darkness for well-watered plants (which itself should be a function of soil water availability, e.g.@misson2004).


Our review of g~min~ of detached leaves shows remarkable plasticity to growing conditions. The lack of clear species trends, when comparing growing conditions, climate of origin, or growth habit, can perhaps be explained by the very large acclimatory potential of g~min~ to growing conditions. We conclude as a result that g~min~ is not a "trait" in the sense that it can be meaningfully summarized for a given species, but has to be linked to growing conditions. Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.


- Also: there is little information on water loss rate of non-leaf surfaces [@oren2001], should be focus of more research


# SuppInfo


E. parramattensis:
We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively. Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%. Inbuilt fans in each chamber blew air gently across the leaf surfaces. Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour. For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~ following @sinclair1986.

Hakeas:
After 8 months of water stress treatment, the epidermal conductance of detached leaves of three plants per species and treatment was estimated by using the rate of water loss. Leaves were collected in the early morning and subsequently recut under distilled water, rehydrated in darkness for 6 h at 4 °C and scanned to determine leaf area. Leaves were then moved to a controlled environment chamber set at 25 °C, 50% relative humidity and low light (~200 umol) and weighed using a precision balance. They were attached to a ‘clothesline’ and allowed to dehydrated slowly. Leaves were weighted every 15 min for 3 h and then every hour for 3h more. For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and gmin from the slope of the linear part of the curve after stomatal closure, normalized by 2-sided leaf area and adjusted for temperature and humidity within the chamber.




# References {.unnumbered}

Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010. "Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)&preview=7 (accessed March 23, 2017).


