---
title: "On the minimum leaf conductance g~0~"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  html_document:
    number_sections: yes
    theme: flatly
csl: global-change-biology.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Belinda Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
**Corresponding author**: 
Remko Duursma

\
E: remkoduursma@gmail.com

\
T: +61(0)45701806



```{r echo=FALSE}
includefigs <- T
library(png)
library(grid)

plotfig <- function(...)grid.raster(readPNG(...))

source("R/load.R")

pval <- function(...)formatPval(..., includeEquality=TRUE)
```


# Introduction: questions about g0.


Prevalence of the Ball-Berry class of stomatal conductance models characterized by dependence on assimilation rate, CO~2~ and VPD or RH.

\
(1)
$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$

where f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), g~1~ the slope parameter, and g~0~ an intercept. The Medlyn model is actually non-linear but can be expressed in an approximate linear form of Eq. 1.

The origin of Eq. 1 is @ball1987 , but no g~0~ was mentioned. It is introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.

@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and also points out that it plays a role in predicting A/g~s~ (and C~i~) at low light [see also @leuning1995].

[@collatz1992] shows that including g0 affects response of gs to RH in the coupled leaf gas exchange model. And "It seems possible that other conditions such as severe drought stress may affect the minimum value of  g  as well, though we currently have no evidence to support this."

g~0~ is often defined as g~s~ when A approaches zero when PPFD approaches zero [@leuning1995] but it is clear from Eq. 1 that g~0~ is also attained when A approaches zero due to other environmental drivers, in particular temperature, and drought stress (which may either reduce photosynthetic capacity or g~1~, or both, see Zhou). It is not clear nor obvious whether these various minima should equal each other.

Though the role and potential for estimation using Eq. 1 is clear, it is not clear whether and how g~0~ compares to other estimates of g~s~ when photosynthesis is zero. It is sometimes assumed that g~0~ must equal g~s~ at night, simply since no photosynthesis occurs at night [@lombardozzi2016; @barnard2013]. Other studies assume that g~0~ equals a 'cuticular conductance` [@manzoni2011; @egea2011], but it has yet to be determined how various minimum conductance estimates compare.

The objective of this paper is to take a closer look at g~0~, its role in modelling fluxes of carbon and water, estimation of g~0~ from regression of Eq. 1 against data. Finally we review the literature for sources of variation in the actual minimum conductance measured in various ways. On the way we try to set straight some common misconceptions about g~0~, and make recommendations for future work. 



# Need for inclusion of g0 in models

@leuning1990 and @leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q. The rate of convergence of ci to ca at low A (low Q) depends on the value of g0; when g0 is large, ci deviates from the asymptotic value at higher values of A than when g0 is small".



[@medlyn2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!

[@hari1986; @hari1999] does not mention it (their Eq 6 assumes g0 = 0)

g~0~ modifies not just conductance in zero light, but also low light and high VPD/ high T. It is commonly observed and makes sense that Ci converges to Ca at zero light, but the coupled leaf gas exchange model does not predict that unless a g0 is included [@leuning1995].

similarly at high T. @kala2016 @urban 
drake in prep
duursma2014


@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'. 

@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance


@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport). However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar). This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.

@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.


\
(2)
$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$

\
where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). 


\
(3)
$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$

where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies. Dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$ [@medlyn2002].

Implemented in the *plantecophys* R package [@duursma2015]


```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

**Figure 1**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~. Parameters were, etc.



We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
Also @gleason2014, @sinclair2000




# Comparison of values for various definitions



**Table 1.** Definitions of minimum conductance

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from a leaf dry-down curve |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |


We compare estimates of g~cut~, g~min~, and g~dark~ from the literature. Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated). We excluded measurements of g~min~ where stomatal leaf sides were sealed, because few data were available and the comparison is not consistent with individual studies where the role of stomata in g~min~ is studied (see SectionXXX).

It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). For g~dark~, we used a recent compilation of measurements from @lombardozzi2017.

For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].

@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.

g~s,minA~ calculated from @lin2015 data. Also g0 fitted using the model of Medlyn et al. 2011, and those from @miner2016 (combined all). 

g~s,low A~ much lower than fitted g~0~ or g~dark~ (see also @dekauwe2015, their Fig. A2).

g~dark~ much higher than g~min~, also confirmed by experiments. In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~. 



```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

**Figure 4**. Comparison of various estimates of minimum conductance. Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable. g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al. 2017), g~0~ : fitted g~0~ in the Medlyn et al. (2011) model with the Lin et al. (2015) database, g~s,low A~ : for the Lin et al. (2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset. Minimum conductance and cuticular conductance are commonly expressed per all surface area but here projected was used because that's more common for g0, gdark.




# Estimation of g~0~ by fitting 


"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]


The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g. @leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2017].


[@barnard2013] mention difficulty of fitting but give no details.

Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.

Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value 

When fitting a Ball-Berry type model, negative values are quite common. A typical approach is to avoid this by setting some value based on guesswork [e.g. @leuning1990], which we assume is quite common when inspecting the many rounded values in the review by 

g~0~ fitted is higher than minimum g~s~ from datasets. If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all. 

Relatively few reports on the plasticity of g~0~, perhaps because it is so poorly estimated. Exceptions:

[@duarte2016] fitted g0 with BB and BBOpti. Lower g0 in heat wave treatment. Also lower gdark in heat wave treatment. But g1 or m not affected by heat waves (again example where a gmin is a responsive parameter!)



```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure3.png")
```

**Figure 3**. Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.






# Nighttime conductance (v. short)

One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero. It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night. Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).

- gdark is variable during the night
- gdark is an active process not just a fixed parameter
- gs very early morning is quite high because in many cases gs increases before the sun comes up
- reviewed extensively elsewhere


Nighttime conductance not necessarily same as minimum conductance
We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]

gnight not constant 
@barbour2007 considerable stomatal control during the night; mirrors daytime
@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.

@misson2004 g0 decreases with pre-dawn water potential

@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
@rescodedios2013 nighttime fluctuation gs independent of environment

@caird2007 For many species, gnight is not stable throughout the night period. Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]

@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]


#  Cuticular conductance (short)

- true cuticular conductance is that of isolated non-stomatal peels, and is very low compared to other definitions
- what is it, and is it variable with anything? yes, temperature, humidity.

Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]

Details of cuticular composition, modes of transport, and composition has been reviewed 
[@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001]

Not really a conductance, since water does not diffuse as a gas through the cuticle. Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle. The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002]. @onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)

gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]


@riederer2001 Review on cuticular conductance. gcut is not correlated with cuticle thickness. No reference to stomata. Cuticles are even less permeable than polymers of equivalent thickness used in packaging.

- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)

# Minimum conductance (longer)

[@hygen1951] early work showing MLD method

Sack et al. 2010 protocol for measurement


- similar to minimum gs observed for the Lin dataset
- plastic trait

One reason gmin is higher than gcut is the role of stomata, even when presumably shut or as shut as they will be.

gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution [@santrucek2004; @brodribb2014]. @muchow1989 correlation stomatal density and gmin suggesting leaky stomata. [@santrucek2004] Direct measurements showing much higher conductance of the leaf side with stomata, though stomata were closed.

@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.

but not always, in @burghardt2003 gmin was the same as cuticular permeance (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).



## Minimum conductance acclimates to growing conditions


@anfodillo2002 gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude (top left page 485).

@herrick1991 gmin increased with altitude in Picea rubens. 

Hypothesis that conifers have low gmin may have to be revisited in the light of @brodribb2014 (check what they actually say!)

@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age

@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1). This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).

@schuster2016 thermal tolerance and gmin relationship across species. 

@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait

[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks. 

- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
page 90 few gmin values from drying curves that we can add.

We present two case studies demonstrating the acclimation of g~min~ to environmental drivers. Both studies use identical methods. The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Leaves were sampled from XXX, n=XXX. The second case study was on *Hakea* plants. Details here (measurement T?). Methods for gmin from drying curves here.




```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**. Analysis of g~min~ from Figure 4 by plant growth form and phylogenetic group. Different letters denote significant differences at $alpha$ = 0.05. The groups 'Angiosperm' and 'Gymnosperm' include trees only. 




```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.



```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure7.png")
```

**Figure 7**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).




# Recommendations and future work


g0 at the moment covers everything, but we need separate ones (night, high VPD/T, drought, low PAR during day)
g0 is poorly estimated from regression - we need to figure out whether it can be measured independently
one option is to use gmin in models, perhaps as averages by growth form / group, but that approach will have to be tested
g0 is certainly not the same as gdark despite the fact that on average they have similar magnitudes. We know for sure gs can be much lower than fitted g0 and gdark in field conditions.






# References {.unnumbered}

Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010. "Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)&preview=7 (accessed March 23, 2017).


