---
title: "On g~0~"
author: ''
date: ''
output:
  html_document:
    theme: flatly
  word_document:
    reference_docx: manuscriptstyle.docx
csl: global-change-biology.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Belinda Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
**Corresponding author**: 
Remko Duursma

\
E: remkoduursma@gmail.com

\
T: +61(0)45701806



```{r echo=FALSE}
includefigs <- F
library(png)
library(grid)

plotfig <- function(...)grid.raster(readPNG(...))

source("R/load.R")

pval <- function(...)formatPval(..., includeEquality=TRUE)
```


# new outline

1. Introduction (short): questions about g0.
2. Need for inclusion of g0 in models
3. Comparison of values for various possible definitions
4. Estimation of g~0~ by fitting 
5. Nighttime conductance (v. short)
6. Cuticular conductance (short)
7. Minimum conductance (longer)
8. Recommendations and future work












# Abstract

# Introduction

Prevalence of the Ball-Berry class of stomatal conductance models characterized by dependence on assimilation rate, CO~2~ and VPD or RH.

\
(1)
$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$

where f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), g~1~ the slope parameter, and g~0~ an intercept. The Medlyn model is actually non-linear but can be expressed in an approximate linear form of Eq. 1.

The g~0~ parameter is the intercept, i.e. g~s~ when A is zero. Depending on the form of f(D), it can also represent g~s~ when D is very large. 

The objective of this paper is to review sources of variation and common misconceptions on the g~0~ parameter.

We argue that

- we need g0 to get responses at low light correct, and probably at high T
- g0 is poorly estimated as the intercept, with the usual regression technique
- can we estimate g0 independently? Some assume that g0 = gdark (also @miner2017 !) but that's rubbish. It is also much larger than cuticular conductance, but seems roughly consistent with gmin (minimum conductance from leaf dry-down curves).
- there is inter-specific variation in gmin that is related to growth form 
- gmin is a plastic trait as it responds to growth at elevated temperature, and growth in dry soil
- models should consider night-time conductance and daily minimum conductance as distinctly separate processes.


g~0~ is often defined as g~s~ when A approaches zero when PPFD approaches zero [@leuning1995] but it is clear from Eq. 1 that g~0~ is also attained when A approaches zero due to other environmental drivers, in particular temperature, and drought stress (which may either reduce photosynthetic capacity or g~1~, or both, see Zhou). It is not clear nor obvious whether these various minima should equal each other.

One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero. It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night. Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).












# Methods

## Data

We compare estimates of g~cut~, g~min~, and g~dark~ from the literature. Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated). It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). For g~dark~, we used a recent compilation of measurements from @lombardozzi2017.

For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].

We present two case studies demonstrating the acclimation of g~min~ to environmental drivers. Both studies use identical methods. The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Leaves were sampled from XXX, n=XXX. The second case study was on *Hakea* plants. Details here (measurement T?). Methods for gmin from drying curves here.


## Model



\
(2)
$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$

\
where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). 


\
(3)
$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$

where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies. Dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$ [@medlyn2002].

Implemented in the *plantecophys* R package [@duursma2015]


# Do we need g~0~ in simulations of water and carbon fluxes?

The first Ball-Berry model did not have a g0 

@leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q. The rate of convergence of ci to ca at low A (low Q) depends on the value of g0; when g0 is large, ci deviates from the asymptotic value at higher values of A than when g0 is small".

[@collatz1992] shows that including g0 affects response of gs to RH in the coupled leaf gas exchange model. And "It seems possible that other conditions such as severe drought stress may affect the minimum value of  g  as well, though we currently have no evidence to support this."


[@medlyn2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!

g~0~ modifies not just conductance in zero light, but also low light and high VPD/ high T. It is commonly observed and makes sense that Ci converges to Ca at zero light, but the coupled leaf gas exchange model does not predict that unless a g0 is included [@leuning1995].

similarly at high T. See also Kala. Figure 1.


@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'. 

@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance

@dekauwe2015 makes this point as well, their Fig. A2 (can we add that also?)

@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport). However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar). This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.

 **Cite Figure 1**


# g~0~ is poorly estimated from regression

"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]


The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g. @leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2017].


Barnard and Bauerle said so too but gave no details.
Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value 

When fitting a Ball-Berry type model, negative values are quite common. A typical approach is to avoid this by setting some value based on guesswork [e.g. @leuning1990], which we assume is quite common when inspecting the many rounded values in the review by 

g~0~ fitted is higher than minimum g~s~ from datasets. If we restric the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all. 


Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]

**Cite Figure 2 and 3**


# Definitions of minimum conductance

- gcut (cuticle without stomata)
- gmin (minimum conductance in situ, stomata closed)
- g0 (minimum conductance in the Ball-Berry-type model)
- gdark (nighttime conductance)

gdark > gmin > gcut
g0?

**Cite Figure 4**

@schuster2016 thermal tolerance and gmin relationship across species. 


@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.

@waldencoleman2013 Very interesting! gdark is different from gmin. Whole-plant WUE correlated with gdark (but not gmin). Lots of references on gdark and gmin and their variation. Closed system to measure gdark+gmin; gmin on detached leaves.



## Cuticular permeance

Not really a conductance, since water does not diffuse as a gas through the cuticle. Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle. The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002]. @onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)

gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]


gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution (santrucek)


## Night-time conductance

Nighttime conductance not necessarily same as minimum conductance
We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]

gnight not constant 
@barbour2007 considerable stomatal control during the night; mirrors daytime
@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.

@misson2004 g0 decreases with pre-dawn water potential

Stomata simply do not close all the way, might stay open regardless of photosynthesis, or are broken. 

@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
@rescodedios2013 nighttime fluctuation gs independent of environment

@caird2007 For many species, gnight is not stable throughout the night period. Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]



@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.

@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]


## Minimum conductance from leaf-dry down curves

Hygen 1951 early work showing MLD method

true cuticular
MLD method (Figure 6) gives higher values but this is probably more useful for models (since these are whole leaves)

santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed. In agreement with @kerstiens1996 who showed 10x higher gmin for stomatal side vs. non-stomatal cuticle.

@riederer2001 Review on cuticular conductance. gcut is not correlated with cuticle thickness. No reference to stomata. Cuticles are even less permeable than polymers of equivalent thickness used in packaging.

@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1). This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).


@burghardt2003 gmin was the same as cuticular permeance (Table 3) (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).


conductance: cuticle + closed stomata + cracks, broken stomata

- Kerstiens
@boyer1997 shows that water but not CO~2~ diffuses across non-stomatal cuticles.

@manzoni2011 assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does. Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut. 

- gmin as a plastic trait. Temperature response generally unresolved [@kerstiens2006]
- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
page 90 few gmin values from drying curves that we can add.

- @burghardt2003 gmin review
- @schreiber1996 cuticle review


[@anfodillo2002] gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude (top left page 485).

[@herrick1991] gmin increased with altitude in Picea rubens. 


[@jordan2007] large increase gmin in the herb Agastachys odorata with leaf age

Sack et al. 2010 protocol for measurement

[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks. 


** cite figures 5,6 and 7**


# Future directions





## Figures


```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

**Figure 1**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~. Parameters were, etc.


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure3.png")
```

**Figure 3**. Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.

```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

**Figure 4**. Comparison of various estimates of minimum conductance. Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable. g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al. 2017), g~0~ : fitted g~0~ in the Medlyn et al. (2011) model with the Lin et al. (2015) database, g~s,low A~ : for the Lin et al. (2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset. Minimum conductance and cuticular conductance are commonly expressed per all surface area but here projected was used because that's more common for g0, gdark.

```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**. Analysis of g~min~ from Figure 4 by plant growth form and phylogenetic group. Different letters denote significant differences at $alpha$ = 0.05. The groups 'Angiosperm' and 'Gymnosperm' include trees only. 

```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.



```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure7.png")
```

**Figure 7**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).



# References {.unnumbered}

Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010. "Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)&preview=7 (accessed March 23, 2017).


