---
title: "On the minimum leaf conductance"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  html_document:
    number_sections: yes
    theme: flatly
csl: global-change-biology.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Belinda Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
**Corresponding author**: 
Remko Duursma

\
E: remkoduursma@gmail.com

\
T: +61(0)45701806



```{r include=FALSE}
includefigs <- F
library(png)
library(grid)

plotfig <- function(...)grid.raster(readPNG(...))

source("R/load.R")

pval <- function(...)formatPval(..., includeEquality=TRUE)
```


# Introduction

\
Stomata exert a strong control on the transpiration rate of leaves. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid dessiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed. We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.

\
Most current-generation process-baed global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.

\
(1)
$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. The effects of photosynthetic photon flux density (PPFD) and leaf temperature are both assumed to enter via the dependence of A~n~ on these drivers. The g~0~ parameter is another definition of the minimum conductance, but we keep the notation g~0~ to indicate that it is a parameter in the Ball-Berry model. Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the inocuous g~0~ parameter has been largely ignored. For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.

\
The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996]. Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951]. This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the dessiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model. 

\
In this review, we aim to compare values of the various estimates of minimum conductance, discuss components and mechanisms of minimum conductance. We place additional emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use. Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq. 1 (and related approaches), and demonstrate why g~0~ needs to be included in models.


# The use of g~0~ in models

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to optimize the rate of photosynthesis per unit water used [@cowan1977]. The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Early work on optimal stomatal conductance models completely ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999], whereas more recent work derives the optimal g~s~ and then adds on a g~0~ [@medlyn2011].

\
In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the dessiccation of leaves in severe drought.

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD). When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

\
(2)
$$A_n/g_s = {C_a}f(D) / g_1$$

and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,

\
(3)
$$C_i / C_a = 1 - \frac{f(D)}{1.6 \cdot g_1}$$

Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.

\
The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~. As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] . It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished). 

\
To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq. 1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,

\
(2)
$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$

\
where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~, 


\
(3)
$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$

\
where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies. The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002]. 

\
Using the coupled leaf gas exchange as implemented in the *plantecophys* R package [@duursma2015], we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1. 


```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

**Figure 1**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~. Parameters were, etc.

\
The third reason for inclusion of a g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants dessiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. We return to measurements of g~min~ from leaf drying curves in Section XXX.


# Comparison of values for various definitions

**Table 1.** Definitions of minimum conductance.

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from a leaf dry-down curve |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |

\
We identify four distinct definitions of minimum conductance, each of which is measured with its own set of methods. Before discussing each of the minimum conductances in detail in the following sections, we here compile literature values to compare the relative magnitude of each of the conductances.

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) and of minimum conductance from leaf dry-down curves (g~min~) were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ re-calculated). We excluded measurements of g~min~ where stomatal leaf sides were sealed, because few data were available and the comparison is not consistent with individual studies where the role of stomata in g~min~ is studied (see SectionXXX). A compilation of night-time (and 'dark') conductance (g~dark~) values was taken from @lombardozzi2017. We also estimated g~0~ from Eq. 1 using the stomatal conductance model of @medlyn2011, using non-linear regression. These estimates were combined with the compilation of g~0~ values from @miner2017, who reported on xxx values from a literature compilation (each of those values was estimated with linear regression of the Ball-Berry model). Finally, we also estimated minimum values of g~s~ attained from a large database of leaf gas exchange data [@lin2015], by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as it represents actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~0~ and g~dark~, all values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].



```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources. 



g~s,low A~ much lower than fitted g~0~ or g~dark~ (see also @dekauwe2015, their Fig. A2).

g~dark~ much higher than g~min~, also confirmed by experiments. In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~. 

@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.



# Estimation of g~0~ by fitting 

The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way. The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative g~0~ estimates (i.e. negative intercepts) are common [@leuning1995; @heroult2013; @miner2017]. In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the  leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]. More recently, @barnard2013 mention the difficulty of fitting Eq. 1 to estimate g~0~ but give no details.

The first problem is that generally in linear regression, the intercept and slope are highly correlated. In Fig. 2, we demonstrate this problem. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, because small errors in g~1~ lead to errors in g~0~. 


Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.


g~0~ fitted is higher than minimum g~s~ from datasets. If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all. 

Relatively few reports on the plasticity of g~0~, perhaps because it is so poorly estimated. Exceptions:

[@duarte2016] fitted g0 with BB and BBOpti. Lower g0 in heat wave treatment. Also lower gdark in heat wave treatment. But g1 or m not affected by heat waves (again example where a gmin is a responsive parameter!)



```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure3.png")
```

**Figure 3**. Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.






# Nighttime conductance (v. short)

One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero. It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night. Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).

- gdark is variable during the night
- gdark is an active process not just a fixed parameter
- gs very early morning is quite high because in many cases gs increases before the sun comes up
- reviewed extensively elsewhere

@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport). However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar). This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.


Nighttime conductance not necessarily same as minimum conductance
We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]

gnight not constant 
@barbour2007 considerable stomatal control during the night; mirrors daytime
@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.

@misson2004 g0 decreases with pre-dawn water potential

@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
@rescodedios2013 nighttime fluctuation gs independent of environment

@caird2007 For many species, gnight is not stable throughout the night period. Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]

@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]


#  Cuticular conductance (short)

- true cuticular conductance is that of isolated non-stomatal peels, and is very low compared to other definitions
- what is it, and is it variable with anything? yes, temperature, humidity.

Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]

Details of cuticular composition, modes of transport, and composition has been reviewed 
[@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001]

Not really a conductance, since water does not diffuse as a gas through the cuticle. Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle. The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002]. @onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)

gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]


@riederer2001 Review on cuticular conductance. gcut is not correlated with cuticle thickness. No reference to stomata. Cuticles are even less permeable than polymers of equivalent thickness used in packaging.

- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)

# Minimum conductance (longer)

[@hygen1951] early work showing MLD method

Sack et al. 2010 protocol for measurement


- similar to minimum gs observed for the Lin dataset
- plastic trait

One reason gmin is higher than gcut is the role of stomata, even when presumably shut or as shut as they will be.

gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution [@santrucek2004; @brodribb2014]. @muchow1989 correlation stomatal density and gmin suggesting leaky stomata. [@santrucek2004] Direct measurements showing much higher conductance of the leaf side with stomata, though stomata were closed.

@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.

but not always, in @burghardt2003 gmin was the same as cuticular permeance (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).



## Minimum conductance acclimates to growing conditions


@anfodillo2002 gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude (top left page 485).

@herrick1991 gmin increased with altitude in Picea rubens. 

Hypothesis that conifers have low gmin may have to be revisited in the light of @brodribb2014 (check what they actually say!)

@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age

@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1). This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).

@schuster2016 thermal tolerance and gmin relationship across species. 

@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait

[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks. 

- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
page 90 few gmin values from drying curves that we can add.

We present two case studies demonstrating the acclimation of g~min~ to environmental drivers. Both studies use identical methods. The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Leaves were sampled from XXX, n=XXX. The second case study was on *Hakea* plants. Details here (measurement T?). Methods for gmin from drying curves here.




```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**. Analysis of g~min~ from Figure 4 by plant growth form and phylogenetic group. Different letters denote significant differences at $alpha$ = 0.05. The groups 'Angiosperm' and 'Gymnosperm' include trees only. 




```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.



```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure7.png")
```

**Figure 7**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).




# Recommendations and future work


g0 at the moment covers everything, but we need separate ones (night, high VPD/T, drought, low PAR during day)
g0 is poorly estimated from regression - we need to figure out whether it can be measured independently
one option is to use gmin in models, perhaps as averages by growth form / group, but that approach will have to be tested
g0 is certainly not the same as gdark despite the fact that on average they have similar magnitudes. We know for sure gs can be much lower than fitted g0 and gdark in field conditions.






# References {.unnumbered}

Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010. "Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)&preview=7 (accessed March 23, 2017).


