---
title: "On the minimum leaf conductance"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  html_document:
    number_sections: yes
    theme: flatly
csl: global-change-biology.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Belinda Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
**Corresponding author**: 
Remko Duursma

\
E: remkoduursma@gmail.com

\
T: +61(0)45701806



```{r include=FALSE}
includefigs <- F
library(png)
library(grid)

plotfig <- function(...)grid.raster(readPNG(...))

source("R/load.R")

pval <- function(...)formatPval(..., includeEquality=TRUE)
```


# Introduction

\

rawson and clarke 1988; gs increased hours pre-dawn in wheat. Nocturnal gs > cuticular (which was 13-50% of total)

@hygen1954 - no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments). "Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"

@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts). 


@bi2017 lower residual transpiration rate in drought-acclimated wheat, in 3 out of 4 genotypes. Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.


@rosasanderson2014, (2013, thesis), lower gmin in drought-stressed peanut



\
Stomata exert a strong control on the transpiration rate of leaves. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed. We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.

\
Most current-generation process-baed global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.

\
(1)
$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (NB they used relative humidity). 

\
The effects of photosynthetic photon flux density (PPFD) and leaf temperature are both assumed to enter via the dependence of A~n~ on these drivers. The g~0~ parameter is another definition of the minimum conductance, but we keep the notation g~0~ to indicate that it is a parameter in the Ball-Berry model. Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the inocuous g~0~ parameter has been largely ignored. For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.

\
The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996]. Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951]. This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the desiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model. 

\
Here we review a) the role and current use of g~0~ in models, b) various definitions of minimum conductance and comparison of literature values, c) sources of variation in minimum conductance. 


# The use of g~0~ in models

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the desiccation of leaves in severe drought. As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD). When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

\
(2)
$$A_n/g_s = {C_a}f(D) / g_1$$

and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,

\
(3)
$$C_i / C_a = 1 - \frac{f(D)}{1.6 \cdot g_1}$$

Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.

\
The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~. As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] . It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished). 

\
To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq. 1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,

\
(2)
$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$

\
where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~, 


\
(3)
$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$

\
where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies. The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002]. 

\
Using the coupled leaf gas exchange as implemented in the *plantecophys* R package [@duursma2015], we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1. 


```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

**Figure 1**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~. Parameters were, etc.

\
The third reason for inclusion of a g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. 
  

# Estimation of g~0~ by fitting 

The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way. The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative g~0~ estimates (i.e. negative intercepts) are common [@leuning1995; @heroult2013; @miner2017]. In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the  leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]. More recently, @barnard2013 mention the difficulty of fitting Eq. 1 to estimate g~0~ but give no details. 

\
It is also telling that there are few reports on the plasticity of g~0~, perhaps because it is so poorly estimated. Exceptions are @duarte2016, who found a lower g0 in a heat wave treatment (although g~1~ was unaffected), and @misson2004 who report on a close negative correlation between pre-dawn leaf water potential and g~0~.

\
Next we demonstrate that the fitting process gives three problems: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated via a bootstrap sampling approach. The correlation between slopes and intercepts is a general aspect of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently. See @becker2007 for more discussion. 


Across datasets, g0 increases when the model fits poorly (Fig. 3 a,b), and the standard error (SE) of g~0~ increases when less variation in a ball-berry coefficient (Fig. 3c). Also, SE is very large compared to the average magnitude of g~0~ across studies (compare Fig. 2 and Fig. 3).


g~0~ fitted is higher than minimum g~s~ from datasets. If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all. 




```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.


```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure3.png")
```

**Figure 3**. Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.



# Comparison of various definitions of minimum conductance

**Table 1.** Definitions of minimum conductance.

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from weight loss of detached leaves |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |


Because estimation of g~0~ is problematic, this begs the question whether suitable estimates of minimum conductance can be arrived at from independent measurements. Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001], or on the side of the leaf that lacks stomata ('adaxial' side). Previous work has thoroughly summarized mechanisms and sources of variation in g~cut~, which will not be repeated in this review. We briefly summarize some important points in SectionXXX.

\
A more common measurement is to detach leaves, and measure leaf weight over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit. Because this method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress, it is perhaps the most promising as an independent measurement of minimum conductance. Because night-time conductance (g~dark~) is frequently used as an independent measurement of g~0~, we compare literature values of g~dark~ with g~min~ and g~cut~, as well as the compilation of fitted values of g~0~ (already shown in Figs X and Y). A final minimum conductance was estimated from the @lin2015 database by extracting actual minimum daytime alues measured, when photosynthetic rates were low.

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996. Technically speaking this is not a conductance, but a 'permeance' because water does not *diffuse* across cuticle [@kerstiens1996]. Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from XXX papers, see Appendix for a list of papers. For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011] (manually filling in numerous gaps), and taxonomic family and order with the taxize R package [@chamberlain2013].



```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources. 


The broad comparison between the various estimates of conductance (FigXX) highlights a number of key features, discussed below.


### Low conductance of isolated cuticles

Literature values of g~cut~ were much lower than any other estimate of conductance, even compared to measurements based on MLD with stomatal side sealed (g~mld,seal~). The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions [@kerstiens1996]. A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, and refer to previous work [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated.

Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002; @schuster2016b]. One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~. Perhaps the answer lies in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).

Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight. 


@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)


Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.



2. g~s,low A~ much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their  Fig. A2).
3. g~dark~ on average eight-fold higher than g~min~. Not many studies measure both g~dark~ and g~min. In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~. @koerner1994 also states that g~s~ in droughted leaves is certainly lower than g~s~ in the dark. 







# Nighttime conductance


A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~). However as our literature comparison showed, g~dark~ is substantially higher compared to g~s~ reached at low photosynthetic rates, and g~min~. Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same species [@howard2007; @caird2007]. The idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments. @rescodedios2016 showed that, in Eucalyptus camaldulensis, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers.







# Minimum conductance (longer)

[@hygen1951] early work showing MLD method


One reason gmin is higher than gcut is the role of stomata, even when presumably shut or as shut as they will be.

gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution [@santrucek2004; @brodribb2014]. @muchow1989 correlation stomatal density and gmin suggesting leaky stomata. [@santrucek2004] Direct measurements showing much higher conductance of the leaf side with stomata, though stomata were closed.

@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.

but not always, in @burghardt2003 gmin was the same as cuticular permeance (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).



## Minimum conductance acclimates to growing conditions


@anfodillo2002 gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude (top left page 485).

@herrick1991 gmin increased with altitude in Picea rubens. 

Hypothesis that conifers have low gmin may have to be revisited in the light of @brodribb2014 (check what they actually say!)

@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age

@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1). This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).

@schuster2016 thermal tolerance and gmin relationship across species. 

@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait

[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks. 

- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.

We present two case studies demonstrating the acclimation of g~min~ to environmental drivers. Both studies use identical methods. The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Leaves were sampled from XXX, n=XXX. The second case study was on *Hakea* plants. Details here (measurement T?). Methods for gmin from drying curves here.

Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time. This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001). Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).





```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**. Analysis of g~min~ from Figure 4 by plant growth form and phylogenetic group. Different letters denote significant differences at $alpha$ = 0.05. The groups 'Angiosperm' and 'Gymnosperm' include trees only. 




```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.




```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure7.png")
```

**Figure 7**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).




# Recommendations and future work


g0 at the moment covers everything, but we need separate ones (night, high VPD/T, drought, low PAR during day)
g0 is poorly estimated from regression - we need to figure out whether it can be measured independently
one option is to use gmin in models, perhaps as averages by growth form / group, but that approach will have to be tested
g0 is certainly not the same as gdark despite the fact that on average they have similar magnitudes. We know for sure gs can be much lower than fitted g0 and gdark in field conditions.






# References {.unnumbered}

Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010. "Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a. cuticular conductance)&preview=7 (accessed March 23, 2017).


